Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  Given more time, we are confident that
this could be completed however.  So far, we have proven the first few lemmas
described in Section~\ref{sec:proof} and defined relations for the operational
semantics and type system as presented in Section~\ref{sec:lang}.  This was much
more difficult than the short natural language proofs would suggest, due to a
combination of the stricter format required by Coq and the number of items that
are elided in the proofs because they were considered trivial or obvious.

We found that the process of mechanizing the lemmas that we proved was a useful
exercise in checking our understanding of our language, statements, and the work
of Simonet and Pottier which we based our approach off of.  We found ourselves
often surprised by the number of items that were elided by the proofs in the
paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
natural language but turned out to be 40 lines of Coq code, not including all of
the supporting definitions and rules.

>>>>

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  As of this writing, we have written definitions of a small-step operational semantics, a big-step operational semantics, typing judgments, and typing derivations for \langName.  We have proven a few lemmas from section~\ref{sec:proof} relating to the operations semantics (including soundness), but remain mired in complexity on the remainder of the mechanization effort.

In the process of mechanizing these lemmas, we empirically observed that the process of translating English into Coq can involve a nearly fifty-fold increase in the size of even the most basic proofs.  While this was partially a natural effect of writing proofs in Coq, we found that much of the complexity stemmed from handling cases elided in the English-Language proofs; and in these cases we found hidden subtleties which appear to have escaped even Simonet and Pottier. We found ourselves
often surprised by the number of items that were elided by the proofs in the
paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
natural language but turned out to be 40 lines of Coq code, not including all of
the supporting definitions and rules.
